---
title: "Arima Assignment"
author: "Sujeeth Shetty"
date: "2/23/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

##wmurders data set from the fpp2 package

```{r loadpackages}
pacman::p_load(fpp2, urca, gridExtra)
theme_set(theme_classic())
```


```{r acf}
#before difference
autoplot(wmurders) + 
  ylab("Total Murdered women") + xlab("Year")
ggAcf(wmurders) 

#After Difference
autoplot(diff(wmurders)) + ylab("Total Murdered women") + xlab("Year")
ggAcf(diff(wmurders))  

#Since there is no seasonality i wont be performing Seasonal difference.
```

**There is trend in the data, hence the data might not be stationary. Also, looking at the ACF plot, it appears we need differencing to reduce trend**


```{r}
##Null Hypothesis the data is stationary
summary(ur.kpss(wmurders))
#test-statistic is higher, we gonna reject null hyp that it is a stationary

# Number of differences needed to make the data stationary
ndiffs(wmurders)

wmurders %>% diff(differences=2) %>% ur.kpss() %>% summary()
#after taking two differences, the series seems to be stationary 


```
**I would not include a constant since we are computing second-order differences.**

```{r ArimaModel}
# Specify (p, d, q) - (1,2,0)

(fit1 <- Arima(wmurders, order = c(1,2,0)))
checkresiduals(fit1)


# Specify (p, d, q) - (2,2,0)

(fit2 <- Arima(wmurders, order = c(2,2,0)) )  
checkresiduals(fit2)

# Specify (p, d, q) - (3,2,0)
(fit3 <- Arima(wmurders, order = c(3,2,0)) )  
checkresiduals(fit3)


```
**I have selected model3 with (p,d,q) value (2,2,0) based on the BIC value & p-value.**
**For the model 3, all lags within ACF plot appear to be within boundaries, but the residuals could be more normal. Overall, I think it is satisfactory.**


```{r forecast}
forecast(fit3,h=3)

autoplot(forecast(fit3,h=3))
```
```{r autoarima}
(fit4<-auto.arima(wmurders))
checkresiduals(fit4)

(fit5<-auto.arima(wmurders, stepwise=FALSE,
           approximation=FALSE, seasonal = FALSE))
checkresiduals(fit5)

```
**auto.arima() didn't give the same model. Even though the BIC value of auto.arima() ARIMA(1,2,1) is lower and all the lags are within boundaries compared to the model I have choosen, the p-value is higher than the 5% significant level. Hence the model isn't significant. So I'll pick my model ARIMA(2,2,0) for forecasting.** 
